---
title: Colors
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Colors}
  %\usepackage[UTF-8]{inputenc}
---


This vignette demonstrates functions in the `colors` range.


```{r, echo=FALSE}
library(shrt)
```

It will be handy to have a vector and a matrix with numbers

```{r}
x = c(-0.2, 0, 0.04, 0.5, 1, NA, 1.2, 0.8)
names(x) = letters[1:length(x)]
x
y = mtrx(seq(-4, 4, length=9), row.names=letters[1:3], col.names=LETTERS[1:3])
y[1,3] = NA
y[2,3] = Inf
y
```




## x2hex	

`x2hex` converts numbers in a [0,1] range into hexadecimal strings that can be used as transparency values

```{r}
x
x2hex(x)
```

The function provides useful behavior in special cases:

 - negative numbers are converted into a two-digit `00`
 - small numbers are always output in two-digit notation with a leading zero
 - `NA` and other non-numbers are preserved as in the original input
 - names associated to the original data are preserved in the output

The function also works when the input is a matrix.

```{r}
xmat = matrix(x, ncol=2)
colnames(xmat) = LETTERS[1:2]
rownames(xmat) = letters[1:4]
xmat
x2hex(xmat)
```

Note the column and row names are preserved under the transformation.




## x2col	

`x2col` converts numbers into colors on a two-color scale. Intermediate colors are created using transparency codes.

```{r}
x
x2col(x)
```

The function provides the following behavior:

 - negative numbers are placed on one color scale (e.g. shades of blue)
 - positive numbers are placed on another scale (e.g. shades of red)
 - the two-color scheme can be changed via argument `col`
 - the numeric values that saturate the color scales can be set by `maxval`
 - `NA` and other non-numbers are preserved 
 - names associated with the input data are preserved
 - values of exactly zero are output as white color. This background color can be changed using argument `bgcol`. This behavior can also be turned off by setting `bgcol` to `NULL`. 

The function also works on matrices

```{r}
y
x2col(y, maxval=3)
```


