# `shrt`: Input and output with files 

This vignette demonstrates functions in the `files` range.


```{r, echo=FALSE}
library(shrt)
```

It will be handy to work with a small table

```{r}
d1 = mtrx(letters[1:8], col.names=LETTERS[1:4], row.names=c("R1", "R2"))
d1
```
	
	
&nbsp;	
## wtht	

`wtht` is a wrapper for `write.table` which by default has `row.names=FALSE`, `col.names=TRUE`, `quote=FALSE`, and `sep='\t'`. The basic usage save a table into a specified file

```{r}
wtht(d1, file="test.1.tsv")
```

`wtht` also provides a way to record row names in a named column. In this mode, the function creates a temporary data structure with a new column, and save this new structure to disk. 

```{r}
wtht(d1, file="test.2.tsv", rowid.column="rowid")
```

Refer to the next section to see the result from these commands.




&nbsp;
## rtht

`rtht` is a wrapper for `read.table` which by default sets `header=TRUE`, `sep='\t'`, and `stringsAsFactors=FALSE`. 

We can use this function to see the files saved above. We can read the first table

```{r}
d2 = rtht("test.1.tsv")
d2
```

Note that row names are absent in this object - they are not recorded in the file. 

The second file we created was meant to contain row names in a new column. We can verify this by reading the file with the same syntax as above

```{r}
d3 = rtht("test.2.tsv")
d3
```

Note here that the table contains a column `rowid` that did not appear in the original dataset. We can optionally eliminate this column whilst reading the data

```{r}
d4 = rtht("test.2.tsv", rowid.column="rowid")
d4
```

This result is now almost equivalent to the orignal data. (Note `d1` was a matrix, but reading a table from disk always returns a data frame.)





```{r, echo=FALSE}
unlink("test.1.tsv")
unlink("test.2.tsv")
```