# `shrt`: Input and output with files 

This vignette demonstrates functions in the `files` range.


```{r, echo=FALSE}
library(shrt)
```

It will be handy to work with a small table

```{r}
d1 = mtrx(letters[1:8], col.names=LETTERS[1:4], row.names=c("R1", "R2"))
d1
```
	



&nbsp;	
## wtht	

`wtht` is a wrapper for `write.table` which by default has `row.names=FALSE`, `col.names=TRUE`, `quote=FALSE`, and `sep='\t'`. The basic usage save a table into a specified file

```{r}
wtht(d1, file="test.1.tsv")
```

`wtht` also provides a way to record row names in a named column. In this mode, the function creates a temporary data structure with a new column, and save this new structure to disk. 

```{r}
wtht(d1, file="test.2.tsv", rowid.column="rowid")
```

Refer to the next section to see the result from these commands.




&nbsp;
## rtht

`rtht` is a wrapper for `read.table` which by default sets `header=TRUE`, `sep='\t'`, and `stringsAsFactors=FALSE`. 

We can use this function to see the files saved above. We can read the first table

```{r}
d2 = rtht("test.1.tsv")
d2
```

Note that row names are absent in this object - they are not recorded in the file. 

The second file we created was meant to contain row names in a new column. We can verify this by reading the file with the same syntax as above

```{r}
d3 = rtht("test.2.tsv")
d3
```

Note here that the table contains a column `rowid` that did not appear in the original dataset. We can optionally eliminate this column whilst reading the data

```{r}
d4 = rtht("test.2.tsv", rowid.column="rowid")
d4
```

This result is now almost equivalent to the orignal data. (Note `d1` was a matrix, but reading a table from disk always returns a data frame.)


```{r, echo=FALSE}
unlink("test.1.tsv")
unlink("test.2.tsv")
```




&nbsp;
## load1

`load1` retrieves data from a file on disk in a similar way to `load`. However, `load1` is specialized to loading single objects from a file. This is handy in situations such as demonstrated below.

Let's save our matrix into an Rda file

```{r}
save(d1, file="test.Rda")
```

The best way to check that the data is really saved would be to load it into
a new R session. Here, we can achieve a similar effect by removing the variable from the environment and then loading it from the file. 

```{r}
rm(d1)
d1
```

As expected, this returned an error. But we can now load the data from the saved file

```{r}
load("test.Rda")
d1
```

A 'feature' of the standard `load` function is that data is associated with the same object names as at the time of `save`. In this case, the matrix appears in the object called `d1`. 

However, this behavior is not always desirable. If we already have an object `d1` in the environment, loading from file would over-write our data

```{r}
d1 = "Important"
d1
load("test.Rda")
d1
```

Thus, we can loose important information through loading of external data. This is where `load1` can be useful

```{r}
d1 = "Important"
d1
dfile = load1("test.Rda")
d1
dfile
```

Thus, we keep all objects in our environment intact and assign the new data into a variable of our choice. 


```{r, echo=FALSE}
unlink("test.Rda")
```



