# `shrt`: Logging and messaging functions

This vignette demonstrate functions in the `messages` range. 

```{r, echo=FALSE}
library(shrt)
msg(action="reset")
```


&nbsp;
## msg

`msg` displays messages on the console and records them for later use. In contrast to many other function in the `shrt` package, this function provides interesting functionality that is not directly provided by base R functions. 

A common way to output a message onto the console is with `cat`

```{r}
cat("hello world")
```

The `msg` function performs a similar role, except that it also outputs a date and time

```{r}
msg("hello")
msg("hello again")
```

An interesting feature is that `msg` internally records all the messages so that they can be retrieved later. 

```{r}
msg(action="show")
```

Note the output here consists of all the previous `msg` messages. It is also possible to retrieve the messages without displaying them

```{r}
mm = msg(action="get")
mm
```

This handy if the messages require further processing, for example saving into a log file. 

When the log is no longer needed, it can be reset 

```{r}
msg(action="reset")
msg(action="show")
```

The output from the last command is not empty. 


&nbsp;
## newmsg

`newmsg` is factory function that creates a new message log. For example, we can create a silent log that does not include the call date. 

```{r}
silent.msg = newmsg(verbose=FALSE, date=FALSE)
class(silent.msg)
```

The new object is a function that works analogously to `msg` above. 

```{r}
silent.msg("hi")
silent.msg("hi again")
```

The messages can again be displayed as above. 

```{r}
silent.msg(action="show")
```

We see all the recorded messages, here without date because of `date=FALSE` in the constructor. 




