---
title: Plot-related 
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Plot-related}
  %\usepackage[UTF-8]{inputenc}
---


This vignette demonstrates functions in the `plot` range.

```{r, echo=FALSE}
library(shrt)
```


&nbsp;

## limq

`limq` provides intervals for defining limits for axes such that the resulting
plot capture data within given quantiles. The basic usage takes in a single
data vector and quantile delimiters.

```{r}
data.100 = seq(0, 100)
limq(data.100, quantiles=c(0.01, 0.99))
```

The resulting range captures data from the 1% quantile to the 99% quantile. In
this case, the result is just the output from function `quantiles`. But `limq`
provides useful functionality through its custom arguments.

Setting argument `lim` can request the output limit to respect a preset lower
or upper limit.

```{r}
limq(data.100, quantiles=c(0.05, 0.95), lim=c(0, NA))
```

In this case, the lower quantile is ignored and the function fills in only the
value of the missing upper limit. 

The function also accepts lists of data and can evaluate liits based on columns
in matrices.

```{r}
data.list = list(M=data.frame(A=0:50, B=-(0:50)),
                 N=data.frame(A=51:100, B=-(51:100)))
limq(data.list, column="A")
limq(data.list, column="B")
```

It is also possible to prettify the output limits via argument `pretty`.

```{r}
limq(seq(3.2, 18.6, by=0.2), quantiles=c(0.05, 0.95), pretty=c(2, 10))
```

In this case, the lower limit is lowered to the nearest multiple of 2. The
upper limit is raised to the nearest multiple of 10. Overall, the interval
includes at least the data in the (0.05, 0.95) interval.


&nbsp;

## widen

Creating a plot often starts by specifying a plot range interval. The function
`widen` starts from an existing interval and calculates a related-but-wider
one. The default action is to update the interval on both sides by 5%.

```{r}
widen(c(0, 100))
```

It is also possible to update the intervals by different amounts on each side,
for example keeping it fixed on the left and expanding by large fraction on the
right,

```{r}
widen(c(0, 100), by=c(0, 0.2))
```

When data is meant to be plotted on a logarithmic scale, the widening of the
axes by a percentage is meant to be perceived after a log transformation. The
`widen` function conveniently can calculate the required numerics,

```{r}
widen(c(1, 100), by=c(0, 0.2), log=TRUE)
```

