# `shrt`: Project management

This vignette demonstrate function in the `projects` range.

```{r, echo=FALSE}
library(shrt)
msg(action="reset")
```


&nbsp;
## initp

`initp` creates a new directory and some subdirectories. 

```{r}
wp = initp("MyProjects")
```

The returned object is a simple list with all the directories.

Note that the messages displayed by `initp` are output via function `msg`. This means they are logged and can be retrieved

```{r}
msg(action="show")
```

See documentation for `msg` for further details. 

```{r, echo=FALSE}
## this block makes sure the directory is cleaned up
unlink("MyProjects", recursive=TRUE)
```



&nbsp;
## reqvars

`reqvars` checks that a set of variable names are defined.

```{r}
a1 = 3
a4 = 0
reqvars(c("a1", "a2", "a3", "a4"))
```

The call prints out messages about variable names that are expected, but not defined. When there are missing variables in the environment, the default action is to call `stop()`. (Argument `halt` can be set to `FALSE` to prevent this behavior.)



