---
title: General-purpose utlity functions
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{General-purpose utility functions}
  %\usepackage[UTF-8]{inputenc}
---


This vignette demonstrates functions in the `utils` range.

```{r, echo=FALSE}
library(shrt)
```




## catv

`catv` is a wrapper for `cat` that prints messages to the console. Unlike the original, this wrapper can be active or silent depending on the value of a verbose setting.

The usual way to control output depending on the value of a `verbose` variable is through an `if` statement, for example,

```{r}
verbose = TRUE
if (verbose) {
   catv("yes, verbose\n")
}
```

The `catv` wrapper accepts an explicit argument to toggle verbose mode,

```{r}
catv("no, not verbose\n", verbose=FALSE)
catv("yes, verbose\n", verbose=TRUE)
```

When the `verbose` argument is not specified, the verbosity status is inferred
from parent environment or from the global options. Thus, it is possible to toggle
verbosity independently of the function call to produce output.

```{r}
verbose=FALSE
catv("this will not appear")
catv("this will not appear either")
verbose=TRUE
catv("Now verbose mode is on, so this will appear\n")
```

Future `catv` statements will also produce output on the console.




## grepf

`grepf` is a search utility looking for patterns within files.
It reads files within a directory and finds patterns within each of them, returning the position of the match within each file.

```{r}
grepf("search", path=".", file.pattern="Rmd$")
```




## grepv

`grepv` is a pattern matching function, a wrapper for `grep`, that returns the actual matching values. For example, let's start with a vector of strings

```{r}
greek = c("alpha", "beta", "gamma", "delta")
```

If we need all items that contain the pattern `a[lm]`, we can use grep,

```{r}
grep("a[lm]", greek, value=T)
```

The `grepv` function syntax is

```{r}
grepv("a[lm]", greek) 
```




## h1, h2, h3

These three functions are wrappers for `head` with preset values for `n`. For example

```{r}
h1(letters)
h2(letters)
h3(letters)
```



## lengrep

`lengrep` is equivalent to `grep` followed by `length`. It returns the number
of hits in a grep query.

```{r}
items = c("foo1", "foo2", "bar")
lengrep("foo", items)
```




## lenu

`lenu` is equivalent to `unique` followed by `length`. It returns the
number of unique elements in a vector.

```{r}
foo = c("f", "o", "o")
lenu(foo)
```




## lst

`lst` creates a new list object and automatically attaches names to its components.

Consider outputing several data objects from a function. The usual way to do that would be to create a named list, often by repeating the components names that are already defined within the function. 

```{r}
a = 1
b = 2
list(a=a, b=b)
```

The `lst` command simplifies the syntax.

```{r}
lst(a, b)
```

The function also takes named arguments, e.g. `lst(a, b, z=3)`.




## mtrx

`mtrx` creates a matrix from input data. Compared to the usual `matrix`, it provides some new possibilities in naming rows and columns.

There are two ways to create a matrix with two rows and four named columns. The first is multi-step

```{r}
mm = matrix(0, nrow=2, ncol=4)
colnames(mm) = greek
mm
```

A second way achieves the same in a single line

```{r}
matrix(0, nrow=2, ncol=4, dimnames=list(rep(NULL, 2), greek))
```

However, this last expression is not very elegant. It requires a null component in the `dimnames` list and the lengths of the row and column names must match the integers specified by `nrow` and `ncol`.

The new `mtrx` function achieves in a slightly more compact form

```{r}
mtrx(0, nrow=2, col.names=greek)
```

Note how the number of columns is automatically determined by the specified names (this is not possible using `matrix`). Besides this usage, `mtrx` also accepts names for rows.





## namesF, namesNA, namesT, namesV

The four functions `namesF`, `namesNA`, `namesT`, and `namesV` are variations on a theme. They act on a vector and provide the names of elements that match particular values. 

Consider a named vector with boolean values

```{r}
foo.boolean = setNames((1:10)%%2==0, letters[1:10])
foo.boolean
```

If we want to know which of letters is associated with a true value, we can match the pattern and fetch the names.

```{r}
names(foo.boolean)[foo.boolean==TRUE]
```

With the `namesT` function, we can instead write

```{r}
namesT(foo.boolean)
```

Note how we avoid writing the vector twice! The function `namesF` behaves similarly but returns the names of elements that are marked as false. 

Consider now that the vector contains numbers or strings.

```{r}
foo.integer = setNames(c(2,4,NA,0,-1,3,NA,7,3,NA), LETTERS[1:10])
foo.integer
```

Suppose we want to know the names that match a certain value, say 3. Finding these is not hard, but it is important to handle NAs carefully. 

```{r}
names(foo.integer)[foo.integer==3]
names(foo.integer)[foo.integer %in% 3]
```

The function `namesV` provides the latter implementation with a simple syntax.

```{r}
namesV(foo.integer, 3)
```

Finally, `namesNA` provides element names that hold NAs.

```{r}
names(foo.integer)[is.na(foo.integer)]
namesNA(foo.integer)
```




## newv

`newv` creates a new vector of specified length, possibly containing names. 

For some data types like integers or characters, we can create a small vector with `rep` and then name the elements with `names`. For example

```{r}
v = rep(0, 4)
names(v) = greek
v
```

The new function can create the same object in a single line. 

```{r}
newv("integer", names=greek)
```

The new function is also handy for creating lists. The traditional way to create a list of a known size is

```{r}
mylist = vector("list", length=4)
setNames(mylist, greek)
```

The new function reproduces this in a single line.

```{r}
newv("list", names=greek)
```

Note that the length of the output object is determined by the names. 




## nlist

`nlist` creates a named list using an existing vector or data frame.

The usual way to create a named list is a two-step process: first create the list using `as.list` and then set names for its elements.

```{r}
mylist = setNames(as.list(letters[1:2]), letters[1:2])
mylist
```

Function `nlist` simplifies this to a single command

```{r}
mylist = nlist(letters[1:2])
mylist
```

The function also allows other uses that start from a data frame or a matrix. That functionality works similarly as `nvec` described below.



## nvec

`nvec` creates a named vector using two columns from a data frame.

The usual way to use matrix/data-frame to create a named vector is a two-step process: first extract values, then assign names.

```{r}
my.data.frame = data.frame(id=c("A", "B"), val=c(2, 6))
myvec = my.data.frame[, "val"]
names(myvec) = my.data.frame[, "id"]
myvec
```

The assignment can be written on one line with `setNames`, but still requires identifying the data frame twice. The `nvec` function is shortcut that requires writing the name of the data frame only once.

```{r}
myvec = nvec(my.data.frame, "val", "id")
myvec
```

The function also checks existence of columns and ensures that the names of the vector are unique.




## p0

`p0` is equivalent to `paste0`. It concatenates its input without adding any spaces or other characters.

```{r}
paste0(greek[1], greek[2])
p0(greek[1], greek[2])
```




## today

`today` returns a current date in a compact character string. 

```{r}
today()
```




## x2df

`x2df` is a conversion function that coerces an input object into a data frame. It is shorthand notations for calls to `data.frame`. For example, let's start with a matrix of characters.

```{r}
charmat = mtrx(letters[1:9], col.names=c("x", "y", "z"), nrow=3)
charmat
```

The conversion produces a related object

```{r}
chardf = x2df(charmat)
chardf
```

The difference from the input is the class.

```{r}
class(chardf)
apply(chardf, 2, class)
```

The default behavior is to convert characters into characters, but it is also possible to set `stringsAsFactors` to true during the function call.

In addition to such simple conversion, the function can also convert vectors into a data frame.

```{r}
myvec = setNames(1:4, letters[1:4])
x2df(myvec)
```


